### 2019-11-16 report/cloud-course-final-report/abstract.tex

The Open Container Initiative (OCI) has been established to create a open standard for container use regardless of the runtime being used to manage the container. However, the OCI only specifies downloading image then unpacking that image into an OCI Runtime filesystem bundle. It does not standardize lifecycle management of the containers, thus each container implements lifecycle functionality in a different manner. It also does not ensure that consistent standards for the security of containers are present.

In this project, we will study the differences in popular runtimes Docker, containerd, and crio. Our focus will be on developing a service focused on ensuring that Center for Internet Security (CIS) Benchmarks for Docker are satisfied across other runtimes to ensure consistent application of security principles irrespective of container runtime differeces. Long term, the goal is to implement a lifecycle mangement solution that enables a common management framework for controlling containers across enviornments. By implementing a service to validate standard secuirty checks across runtimes, we intend to provide a Proof of Concept that such a common lifecycle management is possible.