### 2019-11-16 report/cloud-course-final-report/cis.tex

\subsection*{CIS Benchmarks}
CIS Benchmarks are best practices for the secure configuration of a target system. Available for more than 140 technologies, CIS Benchmarks are developed through a unique consensus-based process comprised of cyber-security professionals and subject matter experts around the world. CIS Benchmarks are the only consensus-based, best-practice security configuration guides both developed and accepted by government, business, industry, and academia.
\input{cis4.tex}
\input{cis5.tex}