### 2019-11-16 report/cloud-course-final-report/docker.tex

\subsubsection*{Docker}
Our primary platform is Docker container run-time. Here, we evaluate the how we implement the CIS checks on Docker. 
First, we find target container's pid from \textit{/run/containerd/io.containerd.runtime.v1.linux/moby/}
\textit{<container-id>/init.pid}}.
In order to issue \textbf{5.1 Do not disable AppArmor Profile} and \textbf{5.2 Verify SELinux security options} benchmarks over Docker, we use procfs and pid. Apparmor and SELinux are security attributes for a given process. So these values are stored under \textit{/proc/<pid>/attr/apparmor/current} and \textit{/proc/<pid>/attr/selinux/current} respectively. By exposing these values, we are able to issue \textbf{5.1} and \textbf{5.2}.

Per container instance, configuration file is created under \textit{/run/containerd/io.containerd.runtime.v1.linux/moby} \textit{/<container-id>/config.json} file. From this file, we get \textbf{cgroup} path of a given container. Exposing this value also corresponds to \textbf{5.24} benchmark.

Further, by using the cgroup path, we are able to access \textit{sysfs} of a given container. \textit{sysfs} is a pseudo file system provided by the Linux kernel that exports information about various kernel subsystems. From this file, we are able to perform \textbf{5.10, 5.11, 5.28}. \textbf{5.10 Memory Limit} of a container is found from: \textit{/sys/fs/cgroup/memory/<container-id>/memory.limit\_in\_bytes}.
\textbf{5.11 CPU Share} of a container is found from: \textit{/sys/fs/cgroup/cpu/<container-id>/cpu.shares}.
\textbf{5.28 PID Limits} of a container is found from: \textit{/sys/fs/cgroup/pids/<container-id>/pids.max}.