### 2019-11-26 report/cloud-course-final-report/abstract.tex

The Open Container Initiative (OCI) has been established to create a open standard for container use regardless of the runtime being used to manage the container. However, the OCI only specifies downloading image then unpacking that image into an OCI Runtime filesystem bundle. It does not standardize lifecycle management of the containers, thus each container implements lifecycle functionality in a different manner. It also does not ensure that consistent standards for the security of containers are present.

In this project, we study the differences in popular runtimes Docker, containerd, and crio. Our focus is on developing an interoperable application focused on ensuring that Center for Internet Security (CIS) Benchmarks are satisfied across runtimes to ensure consistent application of security principles irrespective of container runtime differences. By implementing an application, which can be exposed as a service to validate standard security checks across runtimes, we intend to provide a Proof of Concept that such a common lifecycle management is possible.