### 2019-11-26 report/cloud-course-final-report/crio.tex

\subsubsection*{Cri-o}

Other platform that we evaluate interoperable application is Cri-o  run-time. The process for implementing given benchmarks are pretty similar to Docker and Containerd.
Since different runtimes comes with different configuration files and corresponding structures, we state the implementation details as following.
First, we find target container's pid from its configuration file (i.e., specifically state.json) that is stored under \textit{/var/lib/containers/storage/overlay-containers/} \textit{<container-id>/userdata/state.json}
In order to issue \textbf{5.1 Do not disable AppArmor Profile} and \textbf{5.2 Verify SELinux security options} benchmarks on Containerd, interoperable application
uses procfs and corresponding pid for a container. Apparmor and SELinux are security attributes for a given process. 
These values are stored under \textit{/proc/<pid>/attr/apparmor/current} and \textit{/proc/<pid>/attr/selinux/current} respectively. 
By exposing these values, we are able to issue \textbf{5.1} and \textbf{5.2}.

Per container instance, configuration file is on crio created under \textit{//var/lib/containers/storage/overlay-containers/} \textit{<container-id>/userdata/config.json} file.
From this file, we get \textbf{cgroup} path of a given container. Exposing this value also corresponds to \textbf{5.24} benchmark.

Further, by using the cgroup path, we are able to access \textit{sysfs} and leverage attributes of a container. 
 From this file, we are able to perform \textbf{5.10, 5.11, 5.28}. 
 We specifically use followings:  \textbf{5.10 Memory Limit} of a container is found from: \textit{/sys/fs/cgroup/memory/<container-id>/memory.limit\_in\_bytes}.
\textbf{5.11 CPU Share} of a container is found from: \textit{/sys/fs/cgroup/cpu/<container-id>/cpu.shares}.
\textbf{5.28 PID Limits} of a container is found from: \textit{/sys/fs/cgroup/pids/<container-id>/pids.max}.