### 2019-11-26 report/cloud-course-final-report/design.tex

Interoperable application exists between developer and container runtimes. Our project aims to implement security checks across runtimes. We implement interoperable application in Python language. Right now, this work is a standalone application which is assumed to have access to container configuration files, sysfs, and procfs in the system. The future work for this project is to enable this application as a service which is deployed as daemon that runs on a Kubernetes cluster and ensures CIS benchmarks on container runtimes within that cluster.