### 2019-11-26 report/cloud-course-final-report/intro.tex

Virtualization of resources has emerged in recent decades as a means to run multiple OSes on the same hardware. This particularly serves a useful function as this allows multiple applications to coexist on the same server, enabling efficencies in computing such as server consolidation.

Traditional VMs virtualize hardware resources, which results in the VMs taking up more resources. As such, OS-level virtualization, or Containers, have been developed. By sharing OS resources, containers are lightweight and can be spun up quickly while taking up fewer resources.
Sharing OS resources such as libraries significantly reduces the need to reproduce the operating system code, and means that a server can run multiple workloads with a single operating system installation.
Containers are thus exceptionally light.

Containers are implemented using Linux namespaces and cgroups. \textbf{Namespaces} let you virtualize system resources, like the file system or networking, for each container. \textbf{Cgroups} provide a way to limit the amount of resources like CPU and memory that each container can use.

Docker, introduced in 2013, is a popular runtime to manage containers as it addresses end-to-end management. However, Docker was initially a monolith with features not inherently dependent on each other being bundled together. As a result, alternative runtimes such as CRI-O and contianerd exist which implement container management at varying levels. [1]

The Open Container Initiative (OCI, https://www.opencontainers.org) has been established to create a open standard for container use regardless of the runtime being used to manage the container. However, the OCI only specifies downloading image then unpacking that image into an OCI Runtime filesystem bundle.

It does not standardize lifecycle management of the containers, thus each container implements lifecycle functionality in a different manner. It also does not ensure that consistent standards for the security of containers are present.

In this project, we study the differences in popular runtimes Docker, containerd, and crio. We propose an interoperable application focused on ensuring that Center for Internet Security (CIS) Benchmarks are satisfied across runtimes to ensure consistent application of security principles irrespective of container runtime differences. By implementing an application, which can be exposed as a service to validate standard security checks across runtimes, we intend to provide a Proof of Concept that such a common lifecycle management is possible. Minimum viable product for this project was to implement 5 CIS benchmarks from chapter 5 over at least two different container runtimes. However, we go way beyond the MVP and implement 23 Benchmark checks for docker, 20 for crio, and 16 for containerd. Detailed implementation for the benchmarks can be found in later sections and appendix.

\subsection{Vision and Goals}
Currently if someone wishes to launch an image in a container or perform any other lifecycle management functions on it, they must be sure that the scripts are configured correctly for the target container. For instance, launching images in Docker differs from doing so in crio or containerd. This locks individuals and businesses into whichever container runtime they started with unless they invest the time required to edit the configuration and their scripts which holds the commands for target container.

Our short term goal for this project is to enable the set of Container Runtime tests run in the CIS Docker 1.13.0 Benchmark across any container runtime. These tests are specified in Chapter 5 of the document; example checks include restricting Linux Kernel capabilities within containers, limiting memory usage, and avoiding directly exposing the host devices to the containers. Publishing a minimum viable framework for this purpose will enable users to run their security checks using a single application across the most popular containers.

\subsection*{Users/Personas of the Project}
The intended user is a software developer who is developing, testing or a security engineer who is managing applications and ensuring reliability across containers running on different runtimes.

Example Use Case: A software developer would like to launch an image in CRI-O instead of Docker, because he realizes that CRI-O is more adaptable with Kubernetes, and using this capability will provide this application a lot more scalability. Presently, he needs to deal with changing all the continous-integration scripts in order to be able to test and deploy his application on this new container runtime. With our interoperable framework in place, the developer is at least able to run security checks on the new container runtime with our application by specifying the new target container. In this way, the user's workflow is simplified and can apply a standard security checks across runtimes with minimal effort.

\subsection*{Scope}
The runtimes in scope for capability for this project will be Docker, and CRI-O. containerd is considered a runtime in scope as a stretch goal.

This project aims to ensure that the framework implements commands that satisfy the CIS Docker 1.13.0 Benchmark related to Container Runtimes across our in-scope runtimes. In doing so, users will be enabled to run their security checks with a single application rather than requiring separate suites for each runtime. The MVP is considered to be implementing at least 5 benchmarks in consultation with our mentor. Implementation of the full suite is a stretch goal.

These benchmarks are specified in pp 126-180 of the Benchmark documentation