### 2019-11-26 report/cloud-course-final-report/Article.tex
Add figures

\documentclass[times, twoside, watermark]{zHenriquesLab-StyleBioRxiv}
\usepackage{blindtext}
\usepackage[htt]{hyphenat}
% Please give the surname of the lead author for the running footer
\leadauthor{BU}

\begin{document}

\title{Interoperable}
\shorttitle{Interoperable CIS}

\author[1]{Mert Toslali}
\author[1]{Jonathan Chamberlain}
\author[1]{Felipe Dale Figeman}
\author[1]{Zhengyang Tang}

\affil[1]{BU EC528 - Cloud Computing}

\maketitle

%TC:break Abstract
%the command above serves to have a word count for the abstract
\begin{abstract}
\input{abstract.tex}
\end {abstract}
%TC:break main
%the command above serves to have a word count for the abstract

\section*{Introduction}
\input{intro.tex}

\section*{Design}
\input{design.tex}

\subsection*{Implications and Discussion}

\input{implications.tex}
\input{cis.tex}

\section*{Interoperable Application}

Interoperable application is a Python executable, which accepts two parameters from user. That are, target container runtime and container-id. For instance, to issue CIS Chapter 5 benchmarks over Docker container with id = 0606, user is expected to run command as:
\texttt{./interoperable\_app --docker 0606}.

In this section we demonstrate that we are able to perform more than 10 benchmarks over all container run-times (Docker, Containerd and Cri-o). For the sake of space, we only
mention about some of the benchmark implementation that were more challenging. For the other benchmarks, one can refer to tables that are located in appendix.
\input{docker.tex}
\input{containerd.tex}
\input{crio.tex}

\section*{Other benchmarks}
For the other benchmarks, interoperable application mainly leverages the fields that exist within the corresponding configuration files per container. For example,
to implement \textbf{5.4, 5.7, 5.8, 5.9, 5.12, 5.13, 5.14, 5.15, 5.16, 5.17, 5.18, 5.20 and 5.25} on Docker, we use Hostconfig.json file for a given container. This file
is created and stored per container in : \textit{/var/lib/docker/containers/} \textit{<container-id>/hostconfig.json}. To implement \textbf{5.3, 5.5, 5.21, 5.24} on docker, our application uses config.json under \textit{/run/containerd/io.containerd.runtime.v1.linux/moby/} \textit{<container-id>/config.json}.
\\

For crio, interoperable application implements \textbf{5.3, 5.4, 5.5, 5.7, 5.8, 5.9, 5.12, 5.13, 5.15, 5.16, 5.17, 5.20, 5.21, 5.24} using the config.json file per container. For crio this file is found from \textit{/var/lib/containers/storage/overlay-containers/}\textit{<container-id>/userdata/config.json}
\\

For Containerd, interoperable application implements \textbf{5.3, 5.4, 5.5, 5.17, 5.24, 5.25} using the config.json file per container.
For containerd, this file is found from \textit{/run/containerd/io.containerd.runtime.v2.task/default//} \textit{<container-id>/config.json}.
Interoperable application implements \textbf{5.12, 5.15, 5.16, 5.20, 5.21} using the state.json file per container.
For containerd, this file is found from \textit{/run/containerd/runc/default/}\textit{<container-id>/state.json}
\section*{Appendix}

\begin{figure*}
    \caption{Docker.}
    \centering
      \includegraphics[width=\textwidth,height=10cm]{figures/docker}
  \end{figure*}

  \begin{figure*}
    \caption{CRIO.}
    \centering
      \includegraphics[width=\textwidth,height=10cm]{figures/crio}
  \end{figure*}

  \begin{figure*}
    \caption{Containerd}
    \centering
      \includegraphics[width=\textwidth,height=10cm]{figures/containerd}
  \end{figure*}
\section*{Bibliography}

%% You can use these special %TC: tags to ignore certain parts of the text.
%TC:ignore
%the command above ignores this section for word count
\onecolumn
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Supplementary Information %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\captionsetup*{format=largeformat}
\section{Something about something} \label{note:Note1} 
appendix

%TC:endignore
%the command above ignores this section for word count

\end{document}